PIC simulation of instabilities in relativistic supernovae shocks

### 2021-04-14 papers/Physica2021/Romansky_Physica2021.tex
paper

	\begin{abstract}
                 Modern observations show presence of relativistic supernovae - objects intermediate between regular supernovae and gamma-ray bursts. The typical lorentz-factors of plasma flows in relativistic supernovae are about $1.5$. The process of cosmic ray acceleration is very different from both non-relativistic and ultra-relativistic case. In this paper is presented particle-in-cell modeling of trans-relativistic collisionless shocks. It is shown, that development of magneto-hydrodynamic instabilities, particularly Bell's instability, has strong influence on cosmic ray acceleration. At times larger than $1500$ reverse ion plasma frequency magnetic field, generated by Bell's instability, significantly suppresses the electrons spectrum. This effect should be taken into account for modeling of synchrothron spectrum from relativistic supernovae.
	\end{abstract}

	\section{Introduction}
	Relativistic jets and winds are the generic outflows of the relativistic supernovae \cite{2010Natur.463..513S,2007ApJ...667..351W}, the accreting black holes, both suprmassive in the active galactic nucleis  \cite{1984RvMP...56..255B} as well as the stellar masses in the gamma-ray burst sources, microquasars \cite{2019MmSAI..90...57M,1999PhR...314..575P,2014LNP...876.....R} and pulsar wind nebulae \cite{2019MNRAS.488.5690O,2017SSRv..207..235B,2017JPlPh..83e6301K,2019ApJ...876L...8B}. Relativistic shocks can be formed either inside the flow of low or moderate magnetization (the internal shocks) or can be driven by the outflow colliding with the ambient medium (the external shocks). In many cases the relatistivistic shocks in astrophysical objects are collisionless and are often associated with non-equilibrium particle distributions \cite{2012SSRv..173..309B,2015SSRv..191..519S,2017SSRv..207..319P}. They can accelerate particles to ultra-high energies \cite{2009JCAP...11..009L}. On the one hand, the presence of non-maxwellian components with very long equilibration time requires a kinetic description of the shocked flows. The kinetic structure of relativistic shocks can be modelled within the particle-in-cell approach, which allows one to properly account for the electron scale processes and for the density of the displacement current which is important in relativistic plasma flows.  On the other hand, available computer capabilities do not alow one to provide a fully kinetic modelling of the macroscopic astrophysical flows and, therefore, the fluid MHD description based on such macroscopic parameters as the effective adiabatic index of the flow is widely used. Here we discuss an approach allowing to determine the effective adiabatic indexes from the microscopical particle-in-cell modeling of the mildly-relativistic shocks typical for the relativistic supernovae and gamma-ray burst afterglows.

@@

	Profiles of transversal magnetic field, obtained from numerical simulation, show significant field amplification in precursor of shock wave, see figure \ref{field}. It is result of development of instabilities in plasma, particularly Bell instability \cite{Bell04}, which grows rate is larger then grows rate of other instabilities.

@@

	Spectrum of electrons is shown in figure \ref{spectrum}. One can see, that electrons distribution function at high energies $(\gamma > 300)$ is maximal at times about 1500 inverse protons plasma frequencies and at the longer times spectrum decreases. It may by explained as consequence of magnetic field amplification. Instabilities increase the transverse magnetic field and also the angle between field and shock velocity. As a result shock becomes superluminal \cite{Sironi2011} and can not effectively accelerate particles.

	\begin{figure}[h!]
		\centering
		\includegraphics[width=0.8\textwidth]{fig/spectrum.png} 
		\caption{Evolution of the electrons spectrum in time.}
		\label{spectrum}
	\end{figure}

>>>>

	\begin{abstract}
                 Modern observations show presence of relativistic supernovae - objects intermediate between regular supernovae and gamma-ray bursts. The typical lorentz-factors of plasma flows in relativistic supernovae are about $1.5$. The process of cosmic ray acceleration is very different from both non-relativistic and ultra-relativistic case. In this paper is presented particle-in-cell modeling of trans-relativistic collisionless shocks. It is shown, that development of magneto-hydrodynamic instabilities ii electron-ion relativistic shock, particularly Bell's instability, has strong influence on cosmic ray acceleration. At times larger than $1500$ reverse ion plasma frequency magnetic field, generated by Bell's instability, significantly suppresses the electrons spectrum. This effect should be taken into account for modeling of synchrothron spectrum from relativistic supernovae.
	\end{abstract}

	\section{Introduction}
	Relativistic shocks are widely discussed phenomena in modern astrophysicist. They can be generated in the supernovae \cite{2010Natur.463..513S,2007ApJ...667..351W}, the accreting supermassive black holes \cite{1984RvMP...56..255B}, the stellar masses black holes \cite{2019MmSAI..90...57M,1999PhR...314..575P,2014LNP...876.....R} and pulsar wind nebulae \cite{2019MNRAS.488.5690O,2017SSRv..207..235B,2017JPlPh..83e6301K,2019ApJ...876L...8B}. Relativistic shocks in astrophysical objects are sources of cosmic rays \cite{2012SSRv..173..309B,2015SSRv..191..519S,2017SSRv..207..319P} and can accelerate particles to ultra-high energies \cite{2009JCAP...11..009L}.

	The presence of high energy non-thermal particles leads to development of various magneto-hydrodynamic instabilities. Bell instability \cite{Bell04} can be the most significant due to it's large growth rate. Instabilities modify the flow and has strong influence on process of particle acceleration. Below we present results of particle-in-cell modeling of relativistic shock and show that growth of magnetic field due to instabilities decreases efficiency of electron acceleration on shock front.

@@

	Profiles of transversal magnetic field, obtained from numerical simulation, show significant field amplification in precursor of shock wave, see figure \ref{field}. It is result of development of instabilities in plasma, particularly Bell instability \cite{Bell04}, which growth rate is larger then growth rate of other instabilities.

@@

	Spectrum of electrons is shown in figure \ref{spectrume}. One can see, that electrons distribution function at high energies $(\gamma > 300)$ is maximal at times about 1500 inverse protons plasma frequencies and at the longer times spectrum decreases. It may by explained as consequence of magnetic field amplification. Instabilities increase the transverse magnetic field and also the angle between field and shock velocity. As a result shock becomes quasi-perpendicular and can not effectively accelerate particles \cite{Sironi2011}. On the other hand, protons distribution function constantly increases in time, as shown in figure \ref{spectrump}. We assume, that oscillating transverse field has the strongest influence on the smallest scales, and suppresses he electrons injection in acceleration process.

	\begin{figure}[h!]
		\centering
		\begin{minipage}{0.8\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrum.png} 
			\caption{Temporal evolution of electrons spectrum.}
			\label{spectrume}
		\end{minipage}\hfill
		\begin{minipage}{0.8\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrump.png} 
			\caption{Temporal evolution of protons spectrum.}
			\label{spectrump}
		\end{minipage}
	\end{figure}
